---
title: Hello World!
author: admin
date: '2020-12-15'
categories:
  - planning
tags:
  - blogdown
  - HUGO
  - planning
  - R
slug: post-section-under-construction-plans
lastmod: "`r format(Sys.time(), '%Y-%m-%dT%H:%M:%S%z')`"
featured: no
image:
  caption: vladislav klapin on unsplash
  focal_point: ''
  preview_only: no
output:
  blogdown::html_page:
    toc: true
---


Hey!  I made a web thingamabob!  <span style="color:grey;"> \<polishes fingernails on shoulder\> </span>  Spiffy!

So here's what I have planned thus far for my very own web thingamabob:

# Planned posts:  

1. Building this site with HUGO + wowchemy Academic theme + lots of code-lifting + selectorGadget.  
    + Or "HUGO + wowchemy is a finicky #*!% GAHHH"  
    + Lots of posts on this basic topic, but not many people have written about this __after Academc theme became part of Wowchemy__ and that made things hard to figure out for a first-timer.
    + Also will help me remember what I did and where it goes.
1. Having babies while doing a PhD.  
    + tl;dr: don't do it if you can help it.  Otherwise: coffeeeeeeeeee!
1. Posts on `datagoboop`:
    + Intro to `datagoboop`'s built-in audioplotting functions
    + Intro to basic data $\mapsto$ sound functions in `datagoboop`
    + Building an audioplot from scratch with `datagoboop`
    + Development plans for `datagoboop`
1. Regression with penalized splines 
1. Common penalty contours + P-spline penalty
1. Bayesian model selection basics
1. Parallelizing R scripts on a PC.  
    + don't forget about using source as local job in RStudio!
1. Articles of interest


## Alternate with fun stuff/resources/articles:
- https://www.c82.net/blog/?id=72
- https://seeing-theory.brown.edu/
- http://seeingdata.org/
- https://networkofthrones.wordpress.com/a-primer-on-network-analysis/



# testing .Rmd features

## Divtip 

[desiree de leon's post](https://desiree.rbind.io/post/2019/making-tip-boxes-with-bookdown-and-rmarkdown/)

:::puzzle
puzzle divtip
:::

:::fyi
fyi divtip
:::

:::demo
demo divtip
```{r}
testing <- TRUE
hey <- "it works!"
```
:::


:::note
note divtip
:::


## including plots

```{r chunk_options, echo=F, message=F}
knitr::opts_chunk$set(cache=F, 
                      message=F, 
                      echo=F, 
                      warning=F)
```

### base R
```{r}
x <- 123
plot(x=c(1:x), y = rnorm(x))

```

### ggplot2
```{r}
library(ggplot2)
ggplot(airquality) + 
  geom_point(aes(y=Ozone, x=Solar.R))
```

