---
title: Hello World!
author: admin
date: '2020-12-15'
categories:
  - planning
tags:
  - blogdown
  - HUGO
  - planning
  - R
slug: post-section-under-construction-plans
lastmod: "`r format(Sys.time(), '%Y-%m-%dT%H:%M:%S%z')`"
featured: no
image:
  caption: vladislav klapin on unsplash
  focal_point: ''
  preview_only: no
---


Hey!  I made a web thingamabob!  <span style="color:grey;"> \<polishes fingernails on shoulder\> </span>  Spiffy!

So I just wanted to take a moment to put down what I have planned thus far for my very own web thingamabob.  My plan for this blog is to be a mix of statistics, data vis appreciation, and practical / quality-of-life-type code snippets --- and likely all three at the same time.

Subject to change, of course!




# Table of Contents:

```{r echo=FALSE}
render_toc <- function(
  filename, 
  toc_header_name = "Table of Contents",
  base_level = NULL,
  toc_depth = 3
) {
  # credit Garrick Aden-Buie
  # https://www.garrickadenbuie.com/blog/add-a-generated-table-of-contents-anywhere-in-rmarkdown/
  x <- readLines(filename, warn = FALSE)
  x <- paste(x, collapse = "\n")
  x <- paste0("\n", x, "\n")
  for (i in 5:3) {
    regex_code_fence <- paste0("\n[`]{", i, "}.+?[`]{", i, "}\n")
    x <- gsub(regex_code_fence, "", x)
  }
  x <- strsplit(x, "\n")[[1]]
  x <- x[grepl("^#+", x)]
  if (!is.null(toc_header_name)) 
    x <- x[!grepl(paste0("^#+ ", toc_header_name), x)]
  if (is.null(base_level))
    base_level <- min(sapply(gsub("(#+).+", "\\1", x), nchar))
  start_at_base_level <- FALSE
  x <- sapply(x, function(h) {
    level <- nchar(gsub("(#+).+", "\\1", h)) - base_level
    if (level < 0) {
      stop("Cannot have negative header levels. Problematic header \"", h, '" ',
           "was considered level ", level, ". Please adjust `base_level`.")
    }
    if (level > toc_depth - 1) return("")
    if (!start_at_base_level && level == 0) start_at_base_level <<- TRUE
    if (!start_at_base_level) return("")
    if (grepl("\\{#.+\\}(\\s+)?$", h)) {
      # has special header slug
      header_text <- gsub("#+ (.+)\\s+?\\{.+$", "\\1", h)
      header_slug <- gsub(".+\\{\\s?#([-_.a-zA-Z]+).+", "\\1", h)
    } else {
      header_text <- gsub("#+\\s+?", "", h)
      header_text <- gsub("\\s+?\\{.+\\}\\s*$", "", header_text) # strip { .tabset ... }
      header_text <- gsub("^[^[:alpha:]]*\\s*", "", header_text) # remove up to first alpha char
      header_slug <- paste(strsplit(header_text, " ")[[1]], collapse="-")
      header_slug <- tolower(header_slug)
    }
    paste0(strrep(" ", level * 4), "- [", header_text, "](#", header_slug, ")")
  })
  x <- x[x != ""]
  knitr::asis_output(paste(x, collapse = "\n"))
}

render_toc(filename="index.en.Rmd")
```






# planned posts:  

## stuff I've picked up along the way

1. Building this site with HUGO + wowchemy Academic theme + lots of code-lifting + selectorGadget.  
    + Or "HUGO + wowchemy is a finicky #*!% GAHHH"  
    + Lots of posts on this basic topic, but not many people have written about this __after Academc theme became part of Wowchemy__ and that made things hard to figure out for a first-timer.
    + Also will help me remember what I did and where it goes.
1. Having babies while doing a PhD.  
    + tl;dr: don't do it if you can help it.  Otherwise: coffeeeeeeeeee!
1. Posts on `datagoboop`:
    + Intro to `datagoboop`'s built-in audioplotting functions
    + Intro to basic data $\mapsto$ sound functions in `datagoboop`
    + Building an audioplot from scratch with `datagoboop`
    + Development plans for `datagoboop`
1. Regression with penalized splines 
1. Custom `.rstheme` + colored parens
1. Common penalty contours + P-spline penalty
1. Bayesian model selection basics
1. Parallelizing R scripts on a PC.  
    + don't forget about using source as local job in RStudio!
1. Articles of interest


## other stuff:
- https://www.c82.net/blog/?id=72
- https://seeing-theory.brown.edu/
- http://seeingdata.org/
- https://networkofthrones.wordpress.com/a-primer-on-network-analysis/



# testing .Rmd features

i.e. making sure stuff ain't broke

## divtips

[desiree de leon's post](https://desiree.rbind.io/post/2019/making-tip-boxes-with-bookdown-and-rmarkdown/)

:::puzzle
puzzle divtip
:::

:::fyi
fyi divtip
:::

:::demo
demo divtip
```{r}
testing <- TRUE
hey <- "it works!"
```
:::


:::note
note divtip
:::


## including plots

```{r chunk_options, echo=F, message=F}
knitr::opts_chunk$set(cache=F, 
                      message=F, 
                      echo=F, 
                      warning=F)
```

### base R
```{r}
x <- 123
plot(x=c(1:x), y = rnorm(x))

```

### ggplot2
```{r}
library(ggplot2)
ggplot(airquality) + 
  geom_point(aes(y=Ozone, x=Solar.R))
```

