---
title: Making a Website with Blogdown, R, and Hugo Academic Theme
author: Arnie
date: '2020-12-21'
slug: making-a-website
categories:
  - guide
tags:
  - R
  - blogdown
  - HUGO
subtitle: 'join us on the glorious webtubeways of the blagoblag!'
summary: ''
authors: []
lastmod: '2020-12-21T19:36:17-08:00'
featured: no
image:
  caption: 'https://xkcd.com/181/'
  focal_point: ''
  preview_only: no
projects: []
---

# {-}

I use R, __rmarkdown__, and Github daily, but I started making this website with roughly about exactly zero experience with HTML, Hugo, or the __blogdown__ package, so I wanted to share my experience of getting this site up and running, in the hopes of saving others some pain and time (and documenting for myself).  

There are certainly much better guides out there for the general user; this guide assumes you are starting in a similar place as I did.  I also assume you want basic features of Rmarkdown (generating plots in code chunks, for example) to _just work_, so I will cover what I had to do to make it work.  Also worth mentioning: I'm on a PC, so things might be a bit different for Mac users.

  + [Yihui Xie says figures and such _should just work_](https://bookdown.org/yihui/blogdown/dep-path.html), but they did not for me, I assume because of changes to the Academic Theme.  More on this [below](http://localhost:4321/2020/12/21/making-a-website/#note-on-changes-to-academic-theme).



# Table of Contents:

```{r echo=FALSE}
render_toc <- function(
  filename, 
  toc_header_name = "Table of Contents",
  base_level = NULL,
  toc_depth = 3
) {
  # credit Garrick Aden-Buie
  # https://www.garrickadenbuie.com/blog/add-a-generated-table-of-contents-anywhere-in-rmarkdown/
  x <- readLines(filename, warn = FALSE)
  x <- paste(x, collapse = "\n")
  x <- paste0("\n", x, "\n")
  for (i in 5:3) {
    regex_code_fence <- paste0("\n[`]{", i, "}.+?[`]{", i, "}\n")
    x <- gsub(regex_code_fence, "", x)
  }
  x <- strsplit(x, "\n")[[1]]
  x <- x[grepl("^#+", x)]
  x <- x[!grepl("^# \\{", x)]
  if (!is.null(toc_header_name)) 
    x <- x[!grepl(paste0("^#+ ", toc_header_name), x)]
  if (is.null(base_level))
    base_level <- min(sapply(gsub("(#+).+", "\\1", x), nchar))
  start_at_base_level <- FALSE
  x <- sapply(x, function(h) {
    level <- nchar(gsub("(#+).+", "\\1", h)) - base_level
    if (level < 0) {
      stop("Cannot have negative header levels. Problematic header \"", h, '" ',
           "was considered level ", level, ". Please adjust `base_level`.")
    }
    if (level > toc_depth - 1) return("")
    if (!start_at_base_level && level == 0) start_at_base_level <<- TRUE
    if (!start_at_base_level) return("")
    if (grepl("\\{#.+\\}(\\s+)?$", h)) {
      # has special header slug
      header_text <- gsub("#+ (.+)\\s+?\\{.+$", "\\1", h)
      header_slug <- gsub(".+\\{\\s?#([-_.a-zA-Z]+).+", "\\1", h)
    } else {
      header_text <- gsub("#+\\s+?", "", h)
      header_text <- gsub("\\s+?\\{.+\\}\\s*$", "", header_text) # strip { .tabset ... }
      header_text <- gsub("^[^[:alpha:]]*\\s*", "", header_text) # remove up to first alpha char
      header_slug <- paste(strsplit(header_text, " ")[[1]], collapse="-")
      header_slug <- tolower(header_slug)
    }
    if (header_text != ""){
      paste0(strrep(" ", level * 4), "- [", header_text, "](#", header_slug, ")")
    }
  })
  x <- x[x != ""]
  knitr::asis_output(paste(x, collapse = "\n"))
}

render_toc(filename="index.en.Rmd")
# render_toc(filename)
```



# Preliminaries:

## Note on changes to Academic Theme

Academic Theme for Hugo semi-recently became part of something called Wowchemy, and this appears to have broken a few things for some people, probably because

  1. its folder structure changed, causing uncertainty when following tutorials, and 
  1. how it, um, handles some stuff is like, different.  (I have no idea)

I suspect that these changes are what caused plots not to work for me out of the box.  (Or user error.  Or the fact that my computer is mostly duct tape, [pasta](https://en.wikipedia.org/wiki/Alien_(film)#:~:text=pasta) and a hamster wheel.  But really I think it was the changes to Academic theme.)





## Resources  

(From most used to least used)

  1. [slides by Alison Hill](https://arm.rbind.io/slides/blogdown.html)
    - These are for a different theme

  1. [Wowchemy customization guide](https://wowchemy.com/docs/customization/) 
  
  1. [__blogdown__ documentation by Yihui Xie](https://bookdown.org/yihui/blogdown/)

  
I found HUGO and/or the Academic theme to be pretty finicky, but that may have been at least in part because I started off following the wrong guide.  The guide I wish I had started off following is [this presentation by Alison Hill, given at rstudio::conf 2019](https://arm.rbind.io/slides/blogdown.html).  I say this because I scrapped my first attempt after things were not working, and started again following Alison Hill's slides.





# Building your site:

## prep

+ Install the latest versions of __blogdown__ and Hugo:

```{r eval=FALSE}
if (!requireNamespace("devtools")) install.packages('devtools')
devtools::install_github('rstudio/blogdown')
blogdown::install_hugo()
```


+ make a repo on Github
+ create a new Rproject (in Rstudio: File --> New Project --> )
  - if you don't know how to do this, check out [happygitwithr.com](https://happygitwithr.com/) first.



:::fyi
If you plan to host your website from github, you'll want to create a 2nd repository with the name `<yourGithubusername>.github.io`.  So you can't use that for this step.
:::




## create site with Academic Theme

```{r eval=FALSE}
library(blogdown)
new_site(theme="gcushen/hugo-academic",
         sample=TRUE,
         theme_example=TRUE,
         empty_dirs = TRUE,
         to_yaml = TRUE)
```

+ wait for it to finish, then try using `blogdown::serve_site()` to see what your new website looks like!
  - it's fairly unattractive as-is.  
  - it's also not your website ... yet!



## check deployment

:::fyi
Don't skip this step!  It will help you detect any problems with your setup so far.  I skipped it because I didn't see the point at the time... and ended up rebuilding everything due to a faulty initial download.
:::

Two common options:

- Netlify via Github: best if you want a different domain name, auto-deployment, or a web-form for messages.

- Github: simpler, but a few more steps required for deployment every time you update.


















### folder structure

Your folder structure will look like this (unless they change it yet again!):

```{r eval=FALSE}
├───archetypes
├───assets            # CSS, icons
├───config            # site configuration
│   └───_default
├───content           # play area
│   ├───authors       # your info in the 'admin' folder
│   ├───courses
│   ├───home
│   ├───post
│   ├───project
│   ├───publication
│   ├───slides
│   └───talk
├───data              # themes/fonts go here
├───layouts           # override .html files
├───R
├───resources         # intermediary build files
├───static            # images goes here
│   ├───admin
│   └───media
└───themes            # NO TOUCH!!
```







+ in your .gitignore file, add "public/" and "resources/".  It should look like this:

```{r eval=FALSE}
.Rproj.user
.Rhistory
.RData
.Ruserdata
public/
resources/
```






:::note
The "public" folder contains the actual website.  My understanding is that the "resources" folder contains intermediary files.  The other stuff
:::








# general notes

- wowchemy file structure
- deleting examples
- .rmd support









